---
title: "Project Exploratory"
author: "Jonathan"
output: html_document
---

```{r include=F}
source("Summary.R")
source("Table.R")
source("charts.R")
library(tidyverse)
library(knitr)
library(kableExtra)
```

This report is exploring Police Brutality in Seattle using data from this source: (
http://www.seattle.gov/police-manual/title-8---use-of-force/8050---use-of-force-definitions).
The overall goal is to expose and understand the scale and extent of Police Brutality in Washington
State.

Starting my analysis, this dataset has  total reports, and
 columns. Each row represents a report from a given day, and each column has a different meaning: ID is the identification of the case, Incident_Num is the key for the incident, Incident_Type represents the force that was used:

Level 1 = Type I, wherein the victim undergoes some kind of pain but not significant
or fatal pain due to the force used.

Level 2 = Type II, wherein the victim undergoes some significant pain or injury due to the force used.

Level 3 = Type III, wherein the victim undergoes significant harm & or pain, often
resulting in a loss of Consciousness, great bodily harm, or potentially death.

Level 3 - OIS, Officer Involved Shooting.

The other columns are pretty straightforward: 
Occurred_date_time was when the incident occurred, Precinct is self explanatory, Sector and Beat are police terms for areas within a precinct and then the time and place of the incident within the precinct & sector.

Moving onto some exploration of the data, I was interested in some basic facts first:
Such as the gender of the victims with `r num_male` being male and `r num_female` being female. Next, I made two charts: One depicting the amount of people facing different levels of force seen below:

```{r echo=F}
chart1

```

The trends from this chart are pretty clear, the vast majority of instances where police use force
are not labeled as Fatal or Officer shootings (level 3 & OIS), but there are a disturbing amount of Level 2 cases - it appears roughly half of all total, as a matter of fact. Next, I made a Chart showing the breakdown of each use of force on people by ethnicity:

```{r echo=F}
chart2

```
I was also interested in how many cases had an ethnicity as "unknown", which turns out to be `r race_unspecified` people.

Lastly, I made a chart showing incidents by sector - I used a lolipop format for this one to be more visually interesting, and to understand the frequency a bit better.

```{r echo=F}
chart3
```

While I was exploring this dataset, I was also curious how many police officers were involved from the start of the data on `r date_earliest` to the last date `r most_recent` and found that `r off_num` were involved, with `r rep_off` having more than 1 incident. I also made a table of aggregated table showing the top 10 officers with there total incident counts and ID seen below:

```{r echo=F}
top_10_officers
```
